### 2022-11-27 cleaning_html_data.Rmd
Add files via upload

---
title: "Cleaning HTLM Tags"
author: "Andrew Wells"
date: "11/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Packages
```{r}
library(tidyverse)
library(quanteda)
library(readr)
```

#Loading data
```{r}
df <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/all_data.csv")
```

## Cleaning data
```{r}
#Function for removing HTML tags
cleanFun <- function(htmlString) {
  return(gsub("<.*?>", "", htmlString))
}

#Applying function to clean data
df$summary <- cleanFun(df$summary)
```

#Saving data
```{r}
write.csv(df, "~/Desktop/Semester 5/Classes/4. Text as data/Assignments/Final assignment/all_data_cleaned_summary.csv")
```