### 2022-11-28 Presentation Congressional Data.qmd

---
title: "US Congress Data"
format: revealjs
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished presentation. To learn more about Quarto presentations see <https://quarto.org/docs/presentations/>.

## Bullets

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## Code

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

## Our first steps

ewjfnweögnw

WEFÖGNwegoin

WEGOINW\$GOIErn

```{r}
#our code
```