### 2022-11-29 slides/slides.qmd
data & slides

---
title: "Congress text mining"
title-slide-attributes:
    data-background-image: /layout/background.jpg
    title-slide-style: 
subtitle: "Text as Data Assignment 3"
date: 01.12.2022
format:
    revealjs:
        code-fold: true
        code-tools: true
        theme: sky
        slide-number: true
        preview-links: auto
        footer: "[github.com/juka19/tad_assignment3](https://github.com/juka19/tad_assignment3)"
        transition: slide
        transition-speed: fast
        background-transition: fade
filters:
    - reveal-auto-agenda
    - designmode
auto-agenda:
    bullets: numbered
    heading: Agenda
auto-play-media: true
---

# Research question

# Data acquisition & preprocessing

## Overview
- Data from the 115^th^, 116^th^ and 117^th^ US congress 
- Text summaries of public laws
- Metadata
    - sponsors, cosponsors, actions, policy fields, ...

## Data acquisition

::: {.panel-tabset}

### Challenges 
- Paginated output
- Multiple API calls for each observation
- Request limit of 1000 API calls per hour

[Not knowing about bulk download options of ProRepublica... {{< fa regular face-rolling-eyes >}}]{.fragment}

### Code

```{.python }
@congress_deco(output_format='json')
def get_congress_data(query:str, *args, api_key=os.getenv('US.GOV_API')) -> dict:

        hdr = {
                # specifying requested encoding
                'Cache-Control': 'no-cache',
                'charset': 'UTF-8',
                'X-Api-Key': api_key,
                'User-Agent': random.choice(['Mozilla/5.0', 'Chrome 104.0.0.0', 
                                                'Chrome 52.0.2762.73', 'Chrome 55.0.2919.83'])
                }

        BASE_URL = 'https://api.congress.gov/v3/'

        query_url = BASE_URL + query

        if args: # concatenate variable arguments to url
                query_url += '&' + '&'.join(args)

        return (query_url, hdr)
```

:::
## Data acquisition 3

```{.python }
@congress_deco(output_format='html')
def get_bill_text(query:str) -> str:

        hdr = {
                # specifying requested encoding
                'Cache-Control': 'no-cache',
                'charset': 'UTF-8',
                'User-Agent': random.choice(['Mozilla/5.0', 'Chrome 104.0.0.0', 
                                                'Chrome 52.0.2762.73', 'Chrome 55.0.2919.83'])
                }

        return (query, hdr)
```

## Data

```{.python}
import plotly.ex

```

## Preprocessing

# Methods

# Results

# Discussion

# Conclusion