### 2022-11-30 aw_final_assignment.Rmd
Add files via upload

---
title: "aw_final_assignment"
author: "Andrew Wells"
date: "11/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE}
library(tidyverse)
library(quanteda)
library(quanteda.textstats)
library(ggplot2)
library(gridExtra) 
library(readr)
library(dplyr)
library(quanteda.textplots)
library(uwot)
library(topicmodels)
library(tidytext)
library(ggrepel)
library(vader)
library(tidyr)
library(tidymodels) 
library(textrecipes)
library(plotly)
```

```{r, warning=FALSE}
df <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_11_28.csv")

df$party <- ifelse(df$cosponsor_D_perc > 0.66, "Democrat", ifelse(df$cosponsor_R_perc > 0.66, "Republican", "Both"))
```

## Density of bill cosponsors
```{r, warning=FALSE}

p <- ggplot(df, aes(x = cosponsor_D_perc)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.1, fill="blue") +
  labs(title="Density of bill cosposor party",
       x ="Cosponsor party composition", y = "Density", 
       caption = "Numbers represent proportion of cosponsors from Democratic party, 
       so 0.0 represents bills that were fully Republican and 1.0 represents 
       bills that were fully Democrat.") +
  theme_minimal()

ggplotly(p)
```

```{r, warning=FALSE}
p1 <- df %>%
  filter(party == "Republican") %>%
  ggplot(aes(factor(policy_area))) +
  geom_bar(stat="count", position = "dodge") + 
  scale_fill_brewer(palette = "Set1") + 
    labs(title="Policy areas addressed by Republican bills",
       x ="Count", y = "Policy areas") +
  theme_minimal() +
  coord_flip()

p2 <- df %>%
  filter(party == "Democrat") %>%
  ggplot(aes(factor(policy_area))) +
  geom_bar(stat="count", position = "dodge") + 
  scale_fill_brewer(palette = "Set1") + 
    labs(title="Policy areas addressed by Democrat bills",
       x ="Count", y = "Policy areas") +
  theme_minimal() +
  coord_flip()

grid.arrange(p1, p2, nrow = 1)
```

##Dimensionality plotting

```{r }
corp2 <- corpus(df$summary)

dfmat2 <- corp2 %>%
  tokens(remove_punct = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  dfm() %>%
  dfm_trim(min_termfreq = 5)

embeddings <- umap(as.matrix(dfmat2))

df$x <- embeddings[,1]
df$y <- embeddings[,2]

colordict <- c( "Democrat"="blue","Republican"="red", "Both"="yellow")

p3 <- ggplot(df, aes(x, y, fill=party)) + 
  geom_point(color="grey", shape=21, size=0.5) + 
  scale_fill_manual(values=colordict) +
  theme_bw()

ggplotly(p3)
```

##Sentiment analysis

```{r}
summary_sentiment <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_w_vader.csv")

summary_sentiment$party <- ifelse(summary_sentiment$cosponsor_D_perc > 0.66, "Democrat", ifelse(summary_sentiment$cosponsor_R_perc > 0.66, "Republican", "Both"))

wide_sentiment <- summary_sentiment %>%
  group_by(party, date) %>%
  summarise(score = mean(compound)) %>% 
  pivot_wider(names_from = party, values_from = score) %>%
  select(-c("Both", "NA"))

days <- data.frame(date = seq(as.Date("2017-01-01"),as.Date("2022-12-31"),1))

daily_sentiment <- days %>% 
  left_join(wide_sentiment) %>% 
  pivot_longer(cols = -date, names_to="party", values_to="score")

p4 <- ggplot(daily_sentiment, aes(x=date, y = score, colour=party)) +
  geom_point(aes(y=score), size=1) + 
  theme_minimal() + 
  geom_smooth(method = "loess", se = FALSE)

ggplotly(p4)
```