### 2022-12-01 summary_final_assingment.Rmd
edit_opening

---
title: "Final assignment presentation"
author: "Alexander Kraess, Andrew Wells, Julian Kath & Lorenzo Gini"
date: "11/18/2022"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    hightlight: github
    toc: TRUE
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
library(tidyverse)
library(quanteda)
library(quanteda.textstats)
library(readr)
library(dplyr)
library(quanteda.textplots)
library(uwot)
library(topicmodels)
library(tidytext)
library(ggrepel)
library(vader)
library(tidyr)
library(tidymodels) 
library(textrecipes)
library(plotly)
library(reticulate)
library(igraph)
library(rjson)
library(visNetwork)
```

# Research question

-   Between 2017 and 2022 there have been major shifts in terms of majorities in the US congress:

    -   In the 115th Congress (Jan 2017 - Dec 2018), both chambers were hold by Republicans

    -   In the 116th Congress (Jan 2019 - Dec 2020), the Democrats gained a majority in Congress, the Republicans remained in the majority in the Senate

    -   In the 117th Congress (Jan 2021 - Dec 2022), the Democrats gained Congress and the Senate

-   While one might expect that the post 2016 and 2020 congresses will vary in their policies. However, it is interesting to also focus on the period between 2018 and 2020: When both chambers had different majorities and needed to cooperate.

-   We will focus on the question on whether and how the different majorities had an impact on the policies that have been passed by Congress.

![](https://www.politico.com/dims4/default/0ba7efa/2147483647/strip/true/crop/1160x773+0+0/resize/1290x860!/quality/90/?url=https%253A%252F%252Fstatic.politico.com%252Fa3%252F2c%252Ffbd235e445b4a96e0373e3060e3c%252Fgettyimages-1294351903-1.jpg){width="400"}

## Scraping and cleaning the data

We scraped our data from: <https://data.gov/developers/apis/index.html>

```{r }
df <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_11_28.csv")
df_cosp <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/cosponsors_sponsors.csv")
df_bills <- read_csv("data\\old_data\\all_bills.csv")
head(df, 5)
```

## Creating party variable

```{r }
#if two thirds of the sponsors are democrats, we consider the bill democrat-dominated
#same for republicans
#if there is no clear majority, they are "Both"

df$party <- ifelse(df$cosponsor_D_perc > 0.66, "Democrat", ifelse(df$cosponsor_R_perc > 0.66, "Republican", "Both"))
```

## Summary statistics

## Density of cosponsors

```{r}
ggplot(df, aes(x = cosponsor_D_perc)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") +
  geom_density(alpha=.1, fill="blue") +
  labs(title="Density of bill cosposor party",
       x ="Cosponsor party composition", y = "Density", 
       caption = "Numbers represent proportion of cosponsors from Democratic party, 
       so 0.0 represents bills that were fully Republican and 1.0 represents 
       bills that were fully Democrat.") +
  theme_minimal()
```

## Policy fields

```{r}
fromJSON(file = "slides\\plots\\polar_line_plot") %>% plotly::as_widget()
```

## Creating a corpus

```{r }
df_corp <- df
df_corp <- df_corp %>% rename(text = summary)
corp <- corpus(df_corp)
```

## Creatung a dfm from the corpus

```{r }
dfmat <- corp %>%
  tokens(remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  tokens_replace(pattern = lexicon::hash_lemmas$token, replacement = lexicon::hash_lemmas$lemma) %>%
  tokens_wordstem() %>%
  tokens_remove(c("sec","bill","act", "section", "funds", "shall","must", "use", "author","fund","provid","program","requir","divis","titl","appropri","specifi")) %>%
  dfm()

```

# Wordclouds

## Most common words in all congresses

```{r, warning = FALSE}

dfmatCon <- dfm(corp, remove = stopwords("english"), remove_numbers = TRUE, remove_punct = TRUE, groups = corp$session) %>%dfm_remove(c("sec","bill","act", "section", "funds", "shall","must", "used")) %>%
  dfm_trim(min_termfreq = 3)

textplot_wordcloud(dfmatCon, comparison = TRUE, max_words = 300,
                   color = c("blue", "red"))

#Wordcloud congress 115
```

## Comparing the 115th and 116th congress

```{r, warning = FALSE}

dfmat_115 <- dfm_subset(dfmat, session == 115)
corp_115 <- df %>% filter(session == 115) %>% rename(text = summary) %>% corpus()
modelpart15 <- dfm(corp_115, remove = stopwords("english"),remove_numbers = TRUE, remove_punct = TRUE, groups = corp_115$party) %>%
   dfm_remove(c("sec","bill","act", "section", "funds", "shall","must", "used")) %>%
  dfm_trim(min_termfreq = 3)

mp15 <- textplot_wordcloud(modelpart15, comparison = TRUE, max_words = 300,
                   color = c("green","blue", "red"))

corp_116 <- df %>% filter(session == 116) %>% rename(text = summary) %>% corpus()
modelpart16 <- dfm(corp_116, remove = stopwords("english"), remove_numbers = TRUE, remove_punct = TRUE, groups = corp_116$party) %>% dfm_remove(c("sec","bill","act", "section", "funds", "shall","must", "used")) %>%
  dfm_trim(min_termfreq = 3)

mp16 <- textplot_wordcloud(modelpart16, comparison = TRUE, max_words = 300,
                   color = c("green","blue", "red"))
```

## Dimensionality plotting

```{r }
corp2 <- corpus(df$summary)

dfmat2 <- corp2 %>%
  tokens(remove_punct = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  dfm() %>%
  dfm_trim(min_termfreq = 5)

embeddings <- umap(as.matrix(dfmat2))

df$x <- embeddings[,1]
df$y <- embeddings[,2]

colordict <- c( "Democrat"="blue","Republican"="red", "Both"="yellow")

p <- ggplot(df, aes(x, y, fill=party)) + 
  geom_point(color="grey", shape=21, size=0.5) + 
  scale_fill_manual(values=colordict) +
  theme_bw()

p <- ggplotly(p)
p
```

```{r }
df1 <- df %>%
  mutate(party_full = ifelse(cosponsor_D_perc == 1.0, "Dem",
                             ifelse(cosponsor_R_perc == 1.0, "Rep", NA))) %>%
  drop_na(party_full)

corp3 <- corpus(df1$summary)

dfmat3 <- corp3 %>%
  tokens(remove_punct = TRUE) %>%
  tokens_remove(patter = stopwords("en")) %>%
  dfm() %>%
  dfm_trim(min_termfreq = 5)

embeddings2 <- umap(as.matrix(dfmat3))

df1$x <- embeddings2[,1]
df1$y <- embeddings2[,2]

colordict2 <- c( "Democrat"="blue","Republican"="red")

j <- ggplot(df1, aes(x, y, fill=party)) + 
  geom_point(color="grey", shape=21, size=0.5) + 
  scale_fill_manual(values=colordict2) +
  theme_bw()

j <- ggplotly(j)
j
```

##Sentiment analysis

```{r, warning = FALSE}
summary_sentiment <- read_csv("https://raw.githubusercontent.com/juka19/tad_assignment3/main/data/data_w_vader.csv")
summary_sentiment$party <- ifelse(summary_sentiment$cosponsor_D_perc > 0.66, "Democrat", ifelse(summary_sentiment$cosponsor_R_perc > 0.66, "Republican", "Both"))

wide_sentiment <- summary_sentiment %>%
  group_by(party, date) %>%
  summarise(score = mean(compound)) %>% 
  pivot_wider(names_from = party, values_from = score) %>%
  select(-c("Both", "NA"))
days <- data.frame(date = seq(as.Date("2017-01-01"),as.Date("2022-12-31"),1))
daily_sentiment <- days %>% 
  left_join(wide_sentiment) %>% 
  pivot_longer(cols = -date, names_to="party", values_to="score")
p4 <- ggplot(daily_sentiment, aes(x=date, y = score, colour=party)) +
  geom_point(aes(y=score), size=1) + 
  theme_minimal() + 
  geom_smooth(method = "loess", se = FALSE)+
  scale_color_manual(values = c("blue","red"))
ggplotly(p4)
```

# Network graph of 116 House members

```{r}
df_labels <- df_cosp %>% 
  inner_join(
    df, 
    by = c("number"= "bill number")
  ) %>% 
  inner_join(
    df_bills %>% select(number, originChamber),
    by = 'number'
  ) %>% 
  filter(session == 116, originChamber == 'House') %>% 
  mutate(
    cosponsor_name = str_replace_all(cosponsor_name, '\\w\\.\\s', ''),
    cosponsor_name = str_replace_all(tolower(cosponsor_name), " ", "_"))

df_labels %>% 
  group_by(number) %>% 
  summarize(cosps = paste(cosponsor_name, collapse = " ")) -> df_grouped

corpus(df_grouped$cosps) %>% 
  tokens() %>% 
  fcm() %>% 
  as.matrix() %>% 
  graph_from_adjacency_matrix(mode = "undirected", weighted = T) %>% 
  toVisNetworkData() -> visn

visn$edges$value <- visn$edges$weight

edges <- data.frame(visn$edges) %>% 
  filter(weight > 25)
nodes <- data.frame(visn$nodes) %>% 
  left_join(df_labels %>% select(cosponsor_name, cosponsor_party) %>% distinct(), by = c('label' = 'cosponsor_name')) %>% 
  filter(!duplicated(label)) %>% 
  rename('group' = 'cosponsor_party') %>% 
  mutate(color.background = if_else(group == "R",  "darkred", "darkblue"),
         shadow = T,
         font.color = 'black',
         borderWidth = 2) %>% 
  filter(label %in% edges$from | label %in% edges$to) %>% 
  mutate(label = str_replace_all(label, '_',  ' ') %>% tools::toTitleCase(),
         title = label)

network <- visNetwork(nodes, edges) %>% 
  visNodes(color = list(border = "darkgrey", 
                        borderWidth = 2,
                        highlight = list(border = 'darkgrey', background = 'orange'))
           ) %>% 
  visEdges(color = list(color = '#d3d3d3', highlight = 'orange')) %>% 
  visPhysics(maxVelocity = 3)

network
```

## Topic modeling

```{r}
umap <- read.csv('data\\umap_projection_data.csv')

umap %>% 
  ggplot(aes(x, y, color = cosponsor_D_perc,
             text = paste('Title:' , title, 
                          '<br>Policy Area: ', policyarea, 
                          '<br>Session: ', session, 
                          '<br>Topic: ', cluster))) +  
  geom_point() + 
  scale_color_gradient2(midpoint = .5, name = "Democrat share") + 
  theme_minimal() + 
  labs(x = '', y = '', title = "UMAP Projection of US Congress bill summaries 115-117") +
  theme(panel.background = element_rect(fill='#fbfbfb', colour='#fbfbfb')) -> p

ggp <- ggplotly(p)

ggp
```

## Conclusion

-   In general the legilsation appears quite similar, with exceptions limited to specific policy areas and topics
-   Democratic and Republican cooperation across the different congress sessions is highly dependent on the policy area and topic of the bill
-   There are some key actors that enable working across the aisle and are known for bipartisan collaboration